---
title: "Bluestone Project Donation Report"
author: "Greg Etling"
date: "9/15/2017"
output: pdf_document
---
```{r setup, include = FALSE}
# Knitr defaults
knitr::opts_chunk$set(echo = FALSE,
                      fig.align = 'center', 
                      autodep = TRUE,
                      cache = TRUE)

# Scientific notation defaults
options(scipen=1,digits=4)

require(tidyverse) || {install.packages("tidyverse"); require(tidyverse)}
#require(scales) || {install.packages("scales"); require(scales)}
require(ggthemes) || {install.packages("ggthemes"); require(ggthemes)}
require(RColorBrewer) || {install.packages("RColorBrewer"); require(RColorBrewer)}

# Set base theme
theme_set(theme_minimal())

# Functions
mutate_cond <- function(.data, condition, ..., envir = parent.frame()) {
  condition <- eval(substitute(condition), .data, envir)
  .data[condition, ] <- .data[condition, ] %>% mutate(...)
  .data
}
```

```{r load_data, message = FALSE}
# Load frame
frs <- read_csv("input/initial-fundraise.csv",
         col_types =cols(
           `Order No.` = col_integer(),
           `Funding Date` = col_character(),
           `Payment Method` = col_factor(levels = NULL),
           Appearance = col_factor(levels = NULL),
           Name = col_character(),
           Amount = col_number(),
           `Platform Fee` = col_number(),
           `Perk ID` = col_integer(),
           Perk = col_character())) 

# Rename Columns
frs <- frs %>%
  rename(Order = `Order No.`,
         DateTime = `Funding Date`,
         PaymentType = `Payment Method`,
         Fee = `Platform Fee`,
         PerkID = `Perk ID`)

# Convert DateTime field to POSIXct format
frs$DateTime <- parse_datetime(frs$DateTime, 
                               "%m/%d/%Y %H:%M", 
                               locale = locale(tz = "US/Pacific"))

# Fix Fee percentage values
frs$Fee <- ifelse(frs$Fee >= 1, frs$Fee * .01, frs$Fee)

# Split Perk into Perk and votes
frs <- frs %>% separate(Perk,c("Perk", "votes"), " [+]")
# Convert Perk to ordered factor
frs$Perk <- factor(frs$Perk, levels = c("The Bow of the Head",
                                  "The Acknowledge",
                                  "The Call And Answer",
                                  "The Beacon of Hope",
                                  "The Shining Light"),
                      ordered = TRUE)
# Converting votes to number on currency equivalent scale
frs$votes <- parse_number(gsub(" .*$","",frs$votes)) * 5

```

```{r univariate, eval = FALSE}
require(DescTools) || {install.packages("DescTools"); require(DescTools)}
Desc(frs)
```

# Basic donation statistics
Average donation amount $67
```{r amount1, echo = TRUE}
mean(frs$Amount)
```
Median donation amount $40
```{r amount2, echo = TRUE}
median(frs$Amount)
```

# Beard Vote engagement
41% of people by dollars contributed were engaged by the beard vote.
```{r beardvoteload, message = FALSE}
library(scales)
```
```{r beardvote1, echo = TRUE}
# Sum of non NA votes divided by sum of Amount
percent(sum(frs$votes, na.rm=TRUE)/sum(frs$Amount))
```
37% of people by dollars contributed were engaged by the beard vote.
```{r beardvote2, echo = TRUE}
# Non NA votes divided by total rows
percent(sum(!is.na(frs$votes)) / nrow(frs))
```
Biggest engagement in the beard vote was at the low end and the $250 level:
```{r beardgraph}
ggplot(frs, 
       aes(x = Amount,
           fill = !is.na(votes))) +
  geom_histogram(binwidth = 50) +
  coord_flip() +
  theme_pander() +
  theme(axis.ticks.x = element_blank()) + 
  labs(x = "Donation Amount",
       title = "Donations with Beard Vote Engagement")
```

# Perk engagement
Showing the distribution of donations, colored if they selected a perk:
```{r donationperkgraph}
ggplot(frs, 
       aes(x = Amount,
           fill = Perk)) +
  geom_histogram(binwidth = 50) +
  coord_flip() +
  theme_pander() +
  theme(legend.position = "none",
        axis.ticks.x = element_blank()) + 
  labs(x = "Donation Amount",
       title = "Donations with Selected Perks")
```
Interesting to note in this - perks had lower uptake at lower levels, but definitely had uptake and value in driving the midrange with "The Shining Light" at the $250 level.

Unsurprisingly the top level ($500) is an outlier both in amount and in not being perk-driven.

# Anonymity
There were only 8 Anonymous donors (8.25%):
```{r anon}
sum(frs$Appearance == "Anonymous")
percent(sum(frs$Appearance == "Anonymous") / nrow(frs))
```

Anonymous donors gave larger amounts, but given the small sample it isn't determinative:
```{r anonaov, echo = TRUE}
summary(aov(Amount ~ Appearance,
            data = frs))
```

```{r anonhistgraph}
ggplot(frs, 
       aes(x = Amount,
           fill = Appearance)) +
  geom_histogram(binwidth = 50) +
  coord_flip() +
  theme_pander() +
  theme(axis.ticks.x = element_blank()) + 
  labs(x = "Donation Amount",
       title = "Donations with Anonymity")
```

Even though it isn't statistically significant given the sample size, the average is higher:
```{r donationanongraph}
ggplot(frs, 
       aes(x = Appearance,
           y = Amount,
           fill = Appearance)) +
  geom_boxplot(outlier.shape = NA) +
  coord_flip(ylim = c(0, 200)) +
  theme_pander() + 
  theme(legend.position = "none",
        axis.ticks.y = element_blank()) +
  labs(x = "Anonymous Donor?",
       y = "Donation Amount",
       title = "Average Donation Amount by Anonymity")
```

# Time series - donations over time

There appear to be a couple major clusters of donations, at the beginning and around the start of August:
```{r tsdonations}
ggplot(frs, 
       aes(x = DateTime,
           y = Amount)) +
  geom_point() +
  geom_line() +
  geom_density_2d() + 
  theme_pander() +
  theme(axis.ticks.x = element_blank()) + 
  scale_x_datetime(date_labels = "%b %d",
                   date_breaks = "1 week") +
  labs(x = "Date",
       y = "Donation Amount",
       title = "Donations by date")

ggplot(frs, 
       aes(x = DateTime,
           y = Amount)) +
  geom_bin2d(bins = 15) +
  theme_pander() +
  theme(axis.ticks.x = element_blank()) + 
  scale_x_datetime(date_labels = "%b %d",
                   date_breaks = "1 week") +
  labs(x = "Date",
       y = "Donation Amount",
       title = "Donations by date")

ggplot(frs, 
       aes(x = DateTime,
           y = ..count..,
           fill = DateTime)) +
  geom_histogram() +
  theme_pander() +
  theme(axis.ticks.x = element_blank()) + 
  scale_x_datetime(date_labels = "%b %d",
                   date_breaks = "1 week") +
  labs(x = "Date",
       y = "Number of Donations",
       title = "Number of Donations by date")

```

# TODO
# Couples - split the 'Name' on '&' and create a numPersons feature
# Anonymity - only 8 Anons, slice that a bit more to look for skew
# Would love to add time series of marketing efforts (FB posts, emails, tweets, etc.)
#  Could better pinpoint what may have driven donations
# Knn on Donations over time? Groups of donations by multiple dimensions?

